\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-08 arch.tex
Adapted illustration of model ties to the future to the Shopping example

All of these choices are evaluated sequentially until a consistent
solution with no more flaw for this tick is identified. In order to do
so we need to make the assumption that the current state value of an
internal timeline do not depend on the future or more accurately that
any choices made during this synchronization will not lead to a future
inconsistency (meaning no possible solution) for a future
synchronization. Such assumption deeply impact the set of possible
domains one the reactor can support while remaining complete. Take for
example the following model rule where {\tt Vehicle} is {\em Internal}
and {\tt Command} is {\em External} to the reactor.

\begin{verbatim}
Vehicle::Dive {
  contains(Command.Descend descend);
  descend.start = start + 2;
}
\end{verbatim}

Such model presents the issue that the reactor needs at one point to
take the decision to start the token {\tt Vehicle.Dive} and by doing
so it assumes that in {\em exactly} 2 ticks in the future the {\tt
  Command} timeline will change its state to {\tt Command.Descend}. If
this observation does not occur at the given time this implies that
the reactor estimation was wrong leading him to an inconsistent view
of the world. Conversely the reactor can decide to not switch this
timeline to the {\tt Vehicle.Dive} state and observe 2 ticks in the
future that the {\tt Command.Descend} state value. In this case it
would have missed the opportunity to set its {\tt Vehicle} state to
the correct value which may similarly lead (depending on the rest of
the model) to a similar inconsistency it won't be able to resolve (as
TREX enforces that the past is monotonic).

While this limitation of the system is important to take into account,
it is also important to note that the model snippet we gave here --
while formally acceptable in the general planning problem -- is also
illustrating a bad design for execution in a situated agent. Indeed,
in this model we are tying the current decision to a future outcome
that is {\em External} to the reactor and therefore not necessarily
controllable by this reactor. Therefore we are entailing a part of the
future to past decisions which can lead the system to a dead-end
situation where the model cannot find a correct solution that both do
not conflict with what we claimed in the past and still allow the
planner to find its current {\em Internal} state.

>>>>

All of these choices are evaluated sequentially until a consistent
solution with no more flaw for this tick is identified. In order to do
so we need to make the assumption that the current state value of an
internal timeline do not depend on the future or more accurately that
any choices made during this synchronization will not lead to a future
inconsistency (meaning no possible solution) for a future
synchronization. Such assumption deeply impact the set of possible
domains one the reactor can support while remaining complete. Take for
example this rule from our \texttt{Shopping} model :
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   equals(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11 }
\end{verbatim}
While this model is perfectly standard and acceptable in the overall
case. It can become problematic while tied to the
execution. Specificially it puts a strong tie between the {\em fact} that we
are \texttt{Going} to a location and the {\em expected} future outcome  of
ending at this location. This ties is even stronger due to the
constraint at line 5. Knwoing this aspect assume that the
\texttt{Agent.location} is {\em external} to the reactor which receive
the observation \texttt{Going(Home,SuperMarket)} which synchronization
resolved by producing the observation \texttt{Go(Home, SuperMarket)}
in the Agent timeline. From this point this means that the only
possible outcome on \texttt{Agent.location} is to be
\texttt{At(SuperMarket)} in the directly foreseeable future. In the
real world the fact that we are \texttt{Going} to a certain location
does not necessarly imply that we will succeed on ending up there. An
extreme case could be that due to some exogenous event, the road to
the \texttt{SuperMarket} is closed which could result on an alternate
observation \texttt{At(7th st. \& Foo Ave.)}. This observation
contradicts thre existing plan the reactor maintin and it is important
here to relaize the temp[oral direction of the causal link that is
triggereing this inconsitency. Indeed at the time we will observe this
contradicting \texttt{At}, the rule that generated this conlict is
connected to the past observations \texttt{Going} and \texttt{Go}
which cannot be fully retracted without potential global impact in the
rest of the exisiting reactors -- implying  a huge cost in
backprogation that would be problematic.

More generally tying a token
to future outcomes in a model that is meant to be confronted to rela
executionis something that is problematic. It may be valid ifg you can
gurantee that this will always happen (for example the reactor
managing \texttt{Agent.location} timeline guarantees that the outcome
of the \texttt{Going} will {\em always} result on ending at the traget
location) but should be manipulated while keeping in mind that
predicting the future is not absolute. In this case one could alter
this part of the model as follow in order to give the reactor more
flexibility:
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
The alteration made to relax the constraint on linbe 5 from a strict 
\texttt{equals} to contains and then add the new constraint at line 11
that ensure that the last \texttt{Going} done during this \texttt{Go}
ends up in the location the agent wnats to be. This firts level of
relaxation will therefore allow the agent to usies multiple
\texttt{Going} should the first one fail to reach our objective.  The
probalem can even be relaxed further if needed but this relaxation
gicves already more felxibility for the reactor to recover from local
failures.
